\documentclass[conference]{IEEEtran}

\usepackage{listings}
\usepackage{epsfig}
\usepackage{url}
\usepackage{paralist}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{todonotes}
\usepackage{url}
\usepackage{multirow}
\usepackage{graphicx}

\lstset{
captionpos=b,
language=java,
basicstyle=\scriptsize\ttfamily,
%numbers=left,
%numberstyle=\tiny,
frame=tb,
columns=fullflexible,
showstringspaces=false
}

\begin{document}

\title{The PrimeGame Revolutions 
%\\ {\large A cloud-based collaborative environment for teaching introductory programming}}
}

\author{
	\IEEEauthorblockN{Jens Dietrich}
	\IEEEauthorblockA{Massey University\\Palmerston North,\\ New Zealand\\
	j.b.dietrich@massey.ac.nz}
	\and
	\IEEEauthorblockN{Johannes Tandler}
	\IEEEauthorblockA{Technical University\\ of Dresden\\Dresden, Germany\\
	johannes@jtdl.de}
	\and
	\IEEEauthorblockN{Li Sui}
	\IEEEauthorblockA{Massey University\\Palmerston North,\\ New Zealand\\
		leesui0207@gmail.com}
	\and 
	\IEEEauthorblockN{Manfred Meyer}
	\IEEEauthorblockA{
		Westphalian University of\\ Applied Sciences\\Bocholt, Germany\\
		manfred.meyer@w-hs.de}
}

\maketitle

\begin{abstract}
	
The PrimeGame is an established mathematical programming game that has been used successfully in undergraduate computer science teaching since 2003. To meet the increasing demand for innovative programming tools in undergraduate tertiary and secondary education, we have created SoGaCo, a novel platform to deliver the PrimeGame and similar games to a wide audience via standard web browsers. SoGaCo is designed to have a very low total cost of ownership. This is achieved by enabling teachers to provision a customised collaborative development environment on commodity cloud computing infrastructure. Amongst the unique features of the platform are its social networking features and support for polyglot programming. 

In this paper, we describe the requirements for this system, its design and implementation. We focus on how the scalability and security challenges of an open web-based development environment are addressed. This includes a discussion of the sandboxing and verification techniques we have developed in order to safeguard server-side code execution on the Java Virtual Machine.  
%A preliminary validation suggests that the use of SoGaCo can increase participation and achievement in undergraduate programming courses.

\end{abstract}

\input{introduction.tex}
\input{relatedwork.tex}
\input{design.tex}
\input{security.tex}
\input{scalability.tex}
\input{ui.tex}
\input{conclusion.tex}

\bibliographystyle{plain}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
\bibliography{bibliography}   % name your BibTeX data base


\end{document}
% end of file template.tex

